---
title: "R Notebook"
output:
  html_notebook: default
  html_document: default
---

1. Why R?
R, like SQL, is a language for processing data.  Some tasks are much easier in SQL, other tasks are much easier in R.

| Task                                 | R | SQL |
|--------------------------------------|:-:|:---:|
| Syntax (subjective)                  |   |  x  |
| Reading CSVs                         | x |     |
| ETL / procesing data with many joins |   |  x  |
| Pivoting Data                        | x |     |
| Statistical analysis                 | x |     |
| Charting                             | x |     |
| Fast                                 | x |  x  |
| Very large datasets                  |   |  x  |

An example of an operation that takes much effort in SQL is a [dynamic pivot](https://stackoverflow.com/a/26517282/3827849).

```{r}
# make a dataframe
people <- c("Tom", "John", "Sarah", "Leah")
days <- c("Monday", "Monday", "Tuesday", "Monday")
calories <- c(2100, 1830, 1685, 1900)
cals <- data.frame(people = people, days = days, calories = calories)
head(cals)
```


```{r}
# now lets pivot

# load tidyverse
library("tidyverse")

# pivot - days as columns
spread(cals, days, calories)
```


2. Getting R help

If you know a function name and want to find out what parameters it takes, type `?function` into the console like this:

```{r}
?read.csv
```

If the function name is unknown, Google is your friend.  If all else fails, post of Stack Overflow.

3. Installing and Loading Packages
The default set of functionality that is available in a vanilla R installation is called Base R.  To extend R, you must install packages.

```{r}
# Install the tidyverse package
install.packages("zoo")
```


4. Reading and writing data
    * from CSV/Excel
    
```{r}
# read in a csv
# read_csv is in the tidyverse and dplyr packages
starwars_import <- read_csv("starwars.csv")
glimpse(starwars_import)

# write a csv
write_csv(starwars_import, "starwars_out.csv")
```

    
    
    * from Oracle DW (ROracle & sistools packages)

```{r}
#library(sistools)
#
#oracle_read_table("D_PROGRAM", schema = "DWMART")
```


5. Data Wrangling
    * Base R

Filter rows where name = 'Yoda'
```{r}
#        starwars$name == "Yoda"
starwars[starwars$name == "Yoda",]
```


[dplyr and the tidyverse](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html)
    
Filter rows where name = 'Yoda'
```{r}
filter(starwars, name == "Yoda")

# or...

starwars %>% filter(name == "Yoda")
```

6. Plotting with ggplot2

**Histogram**
```{r}
ggplot(data = starwars, aes(x=height)) + geom_histogram()
```

**Scatter**
```{r}
#ggplot(data = starwars, aes(x = height, y = mass)) +
ggplot(data = starwars, aes(x = height, y = mass, colour = gender)) +
  geom_point(size = 3)
```


7. Rstudio tips
    * Projects
    * Snippets
